---
title: "Assignment B1"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE
)

# Load necessary libraries
library(dplyr)
library(testthat)
```

## Exercise 1: Make a Function

I will create a function called `summarize_data` that takes a data frame, groups it by a specified column, and then summarizes a numeric column by calculating the mean and standard deviation. This function will handle missing values appropriately and ensure flexibility in its inputs.

In the function `summarize_data`, the parameters are named to clearly reflect their roles within the function:

- `data`: This parameter is named `data` to indicate that it should be a dataframe, clearly specifying that this is the dataset on which the summarization operations will be performed.
- `group_var`: The name `group_var` is used to denote the variable within `data` that defines the groups for aggregation. This name makes it clear that this parameter should be the name of a column used to group the data.
- `summary_var`: Similarly, `summary_var` explicitly indicates that this parameter should be the column name of the data that will be summarized, making it intuitive for users to understand that this is the variable on which summary statistics like mean and standard deviation will be calculated.
- `na.rm`: The `na.rm` parameter stands for "NA remove" and is a common naming convention in R, which clearly communicates its purpose to control the handling of NA (missing) values in the data summarization process. Setting it to `TRUE` by default indicates that NA values will be removed before computing the summaries, which is a typical requirement for statistical calculations to avoid biased or erroneous results.

```{r create-function}
#' Summarize Data by Group
#'
#' This function takes a data frame, groups it by a specified column,
#' and summarizes a numeric column by calculating the mean and standard deviation.
#'
#' @param data A data frame to be summarized.
#' @param group_var A string specifying the column name to group by.
#' @param summary_var A string specifying the numeric column to summarize.
#' @param na.rm A logical value indicating whether to remove NA values. Defaults to TRUE.
#'
#' @return A data frame with the grouping variable and the calculated mean and standard deviation.
#'
#' @examples
#' summarize_data(mtcars, "cyl", "mpg")
summarize_data <- function(data, group_var, summary_var, na.rm = TRUE) {
  # Check if group_var and summary_var are in the data frame
  if (!all(c(group_var, summary_var) %in% names(data))) {
    stop("Specified group_var or summary_var does not exist in the data frame.")
  }
  
  # Ensure that summary_var is numeric
  if (!is.numeric(data[[summary_var]])) {
    stop("summary_var must be a numeric column.")
  }
  
  # Process 'data' to group and summarize specified variables
  data %>%
    # Group data by the specified 'group_var' column
    group_by(.data[[group_var]]) %>%
    # Calculate summary statistics for the 'summary_var' column:
    # 'mean' - computes mean, ignoring NA values if 'na.rm' is TRUE
    # 'sd' - computes standard deviation, also ignoring NAs based on 'na.rm'
    summarise(
      mean = mean(.data[[summary_var]], na.rm = na.rm),
      sd = sd(.data[[summary_var]], na.rm = na.rm)
    ) %>%
    # Remove grouping structure from the data to return a regular data frame
    ungroup()
}
```

## Exercise 2: Document your Function

The documentation for `summarize_data` is included within the same code chunk using `roxygen2` comments.

## Exercise 3: Include Examples

I will demonstrate how to use the `summarize_data` function with a few examples.

### Example 1: Summarizing Miles Per Gallon by Cylinder Count in `mtcars`

```{r example-1}
# Summarize mpg by cyl in mtcars
summary_mtcars <- summarize_data(mtcars, "cyl", "mpg")
print(summary_mtcars)
```

### Example 2: Handling NA Values

I will introduce some `NA` values into the `mtcars` dataset to see how the function handles them.

```{r example-2}
# Introduce NA values
mtcars_na <- mtcars
mtcars_na$mpg[c(1, 5, 10)] <- NA

# Summarize with na.rm = TRUE
summary_na_rm <- summarize_data(mtcars_na, "cyl", "mpg", na.rm = TRUE)
print(summary_na_rm)

# Summarize with na.rm = FALSE
summary_na_not_rm <- summarize_data(mtcars_na, "cyl", "mpg", na.rm = FALSE)
print(summary_na_not_rm)
```

## Exercise 4: Test the Function

Formal tests are essential to ensure that the function behaves as expected under various conditions. I will use the `testthat` package to write these tests.

```{r tests}
test_that("summarize_data works correctly with no NAs", {
  result <- summarize_data(mtcars, "cyl", "mpg")  # Perform the operation
  expect_true(is.data.frame(result))  # Validate the result is a dataframe
  expect_equal(nrow(result), length(unique(mtcars$cyl)))  # Compare row count to unique groups
  expect_true(all(c("mean", "sd") %in% names(result)))  # Ensure 'mean' and 'sd' are in column names
})

test_that("summarize_data handles NA values correctly when na.rm = TRUE", {
  mtcars_na <- mtcars
  mtcars_na$mpg[1:3] <- NA  # Introduce NA values into the dataset
  result <- summarize_data(mtcars_na, "cyl", "mpg", na.rm = TRUE)  # Summarize data with NA removal
  expect_true(all(!is.na(result$mean)))  # Check for absence of NAs in 'mean'
  expect_true(all(!is.na(result$sd)))  # Check for absence of NAs in 'sd'
})

test_that("summarize_data handles NA values correctly when na.rm = FALSE", {
  mtcars_na <- mtcars
  mtcars_na$mpg[1:3] <- NA  # Introduce NA values into the dataset
  result <- summarize_data(mtcars_na, "cyl", "mpg", na.rm = FALSE)  # Summarize data without removing NAs
  expect_true(any(is.na(result$mean)))  # Expect NAs to be present in 'mean'
  expect_true(any(is.na(result$sd)))  # Expect NAs to be present in 'sd'
})

test_that("summarize_data throws an error for non-numeric summary_var", {
  mtcars_test <- mtcars %>%
    mutate(gear_factor = as.factor(gear))  # Convert 'gear' to a factor to create a non-numeric summary variable
  expect_error(summarize_data(mtcars_test, "cyl", "gear_factor"))  # Expect error due to non-numeric data
})

test_that("summarize_data throws an error when group_var does not exist", {
  expect_error(summarize_data(mtcars, "nonexistent_var", "mpg"))  # Expect error due to non-existent group variable
})
```
